---
title: "EC-Earth3_ssp245 annual, local tos to global tas pattern scaling"
author: "ACS"
date: "12/19/2019"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(RColorBrewer)


devtools::load_all(path = '~/Documents/GitHub/fldgen')

knitr::opts_chunk$set(echo = TRUE)
```


# Setup {.tabset}

## Select model + helpful strings and other files
```{r names}
esmname <- 'rgr2p5-EC-Earth3_ssp245'
esmfilelist <- list.files(paste0('cmip6_netcdf/', substr(esmname, 8, nchar(esmname))), full.names = TRUE)
esmfilelist <- esmfilelist[grepl('rgr2p5-', esmfilelist)]

outfileyrs <- 2000:2100

# CMIP6 offers gr (regridded) and gn (native) resolutions; cannot get gn to open 
# in some cases, there are gr1, gr2, etc. Unclear the details/differences. 
# probably regridded to different scale. Fine as long as consistent within
# a given variable (tos vs tas)
gridchoice <- 'gr' 


coralsites <- as_tibble(read.csv('../data/cell_site_lat_longs.csv'))

hawaii <- paste0(rep('hi', 15), 1:15)
florida <- paste0(rep('fl', 9), 1:9)
pr <- paste0(rep('pr_usvi', 8), 1:8)
```


## Helper function to pull off, shape, and save  a year's data that can be read by COMBO
```{r helperfunction}

# helper function to pull off, shape, and save  a year's data
make_single_year_output <- function(outdir, tosdata, tgavdata, year){
  
  ind <- which(tgavdata$year == year)
  
  tibble(tos = tosdata[ind,]) %>%
    mutate(colnumber = as.numeric(row.names(.))) %>%
    left_join(lonlat, by = 'colnumber') %>%
    select(-colnumber) %>% 
    mutate(tos = round(tos, digits = 2)) %>%
    complete(gridlat = fulllat, gridlon = fulllon) %>%
    replace_na(list(tos = -999)) %>%
    spread(gridlon, tos) %>%
    arrange(-gridlat) %>%
    # numeric column names so matlab understands
    rename( `1` = gridlat) %>% 
    mutate(`2` = `1`) -> 
    # rename( `degN/degE` = gridlat) %>%
    # mutate(`: degN` = `degN/degE`) -> 
    annualdata
  
  annualdata <- rbind(annualdata, names(annualdata))
  
  # csv with numeric column names and no header helps matlab
  write.table(annualdata,
              paste0('pscl_out_files/COMBOgrid/', outdir, '/', year, '_', esmname, '_hector_', outdir, '.csv'),
              sep = ",", quote = FALSE, row.names=FALSE)
  
}
```

# Load the pattern
```{r pattern load}

# the pattern is only for the ocean grid cells.
# To make the .OUT files for combo, the full 
# grid is necessary. 
# It's also convenient  to have the grid 
# lon and lat values over the ocean and their
# corresponding colnumber in the `tos` data 
# that has no NAs.
# So these are included in the rds file for 
# each pattern. 

x <- readRDS(paste0('patterns/', esmname, '_annualtos.rds'))

annual_pattern <- x$annual_pattern
lonlat <- x$lonlat
fulllon <- x$fulllon
fulllat <- x$fulllat

rm(x)
```

# Load and process hector data as needed
Beacuse the ESM `tas` data used to calculate `tgav` and form the pattern is in terms of actual degrees, not anomaly from pre-industrial, Hector temperatures MUST be converted to actual temperature via adding 288.15.


```{r hectorload}

# hector rcp4.5 tgav trajectory
as_tibble(read.csv('hectordata/hector_temp_rcp45.csv', stringsAsFactors = F)) %>%
  # convert from delta tgav to tgav and then to celsius %>%
  mutate(temp = temp + 288.15 - 273.15) ->
  hector.tgav45


# hector rcp2.6 tgav trajectory
as_tibble(read.csv('hectordata/hector_temp_rcp26.csv', stringsAsFactors = F, comment.char = '#')) %>%
  # convert from delta tgav to tgav and then to celsius %>%
  mutate(temp = temp + 288.15 - 273.15) ->
  hector.tgav26


# hector noparis tgav trajectory
as_tibble(read.csv('hectordata/temp_noparis.csv', stringsAsFactors = F, comment.char = '#')) %>%
  # convert from delta tgav to tgav and then to celsius %>%
  mutate(temp = tgav + 288.15 - 273.15) ->
  hector.tgavnoparis


# hector paris tgav trajectory
as_tibble(read.csv('hectordata/temp_paris.csv', stringsAsFactors = F, comment.char = '#')) %>%
  # convert from delta tgav to tgav and then to celsius %>%
  mutate(temp = tgav + 288.15 - 273.15) ->
  hector.tgavparis


# hector parisplus tgav trajectory
as_tibble(read.csv('hectordata/temp_parisplus.csv', stringsAsFactors = F, comment.char = '#')) %>%
  # convert from delta tgav to tgav and then to celsius %>%
  mutate(temp = tgav + 288.15 - 273.15) ->
  hector.tgavparisplus


# hector parisref tgav trajectory
as_tibble(read.csv('hectordata/temp_parisref.csv', stringsAsFactors = F, comment.char = '#')) %>%
  # convert from delta tgav to tgav and then to celsius %>%
  mutate(temp = tgav + 288.15 - 273.15) ->
  hector.tgavparisref


```

# Construct the out files

## Hector rcp4.5 with the pattern


```{r}

# Get the annual TOS in each grid cell as predicted 
# by the annual pattern for the Hector tgav 
hector_annual_tos <- pscl_apply(annual_pattern, as.vector(hector.tgav45$temp))

# make the outfile for every year
invisible(
  for(yr in outfileyrs) {
    
    make_single_year_output(outdir = 'RCP45', tgavdata = hector.tgav45, 
                            tosdata = hector_annual_tos, year = yr)
    
  }
)

```


## Hector rcp2.6 with the pattern
```{r}
# by the annual pattern for the Hector tgav
hector_annual_tos26 <- pscl_apply(annual_pattern, as.vector(hector.tgav26$temp))

# make the outfile for every year
invisible(
  for(yr in outfileyrs) {
    
    make_single_year_output(outdir = 'RCP26', tgavdata = hector.tgav26, 
                            tosdata = hector_annual_tos26, year = yr)
    
  }
)
```


## Hector no paris
```{r}
# by the annual pattern for the Hector tgav
hector_annual_tosnoparis <- pscl_apply(annual_pattern, as.vector(hector.tgavnoparis$temp))

# make the outfile for every year
invisible(
  for(yr in outfileyrs) {
    
    make_single_year_output(outdir = 'noparis', tgavdata = hector.tgavnoparis, 
                            tosdata = hector_annual_tosnoparis, year = yr)
    
  }
)

```


## Hector paris
```{r}
# by the annual pattern for the Hector tgav
hector_annual_tosparis <- pscl_apply(annual_pattern, as.vector(hector.tgavparis$temp))

# make the outfile for every year
invisible(
  for(yr in outfileyrs) {
    
    make_single_year_output(outdir = 'paris', tgavdata = hector.tgavparis, 
                            tosdata = hector_annual_tosparis, year = yr)
    
  }
)

```


## Hector paris plus
```{r}
# by the annual pattern for the Hector tgav
hector_annual_tosparisplus <- pscl_apply(annual_pattern, as.vector(hector.tgavparisplus$temp))

# make the outfile for every year
invisible(
  for(yr in outfileyrs) {
    
    make_single_year_output(outdir = 'parisplus', tgavdata = hector.tgavparisplus, 
                            tosdata = hector_annual_tosparisplus, year = yr)
    
  }
)

```


## Hector paris ref
```{r}
# by the annual pattern for the Hector tgav
hector_annual_tosparisref <- pscl_apply(annual_pattern, as.vector(hector.tgavparisref$temp))

# make the outfile for every year
invisible(
  for(yr in outfileyrs) {
    
    make_single_year_output(outdir = 'parisref', tgavdata = hector.tgavparisref, 
                            tosdata = hector_annual_tosparisref, year = yr)
    
  }
)

```