---
title: "GFDL-CM4_ssp245 annual, local tos to global tas pattern scaling"
author: "ACS"
date: "12/19/2019"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(RColorBrewer)


devtools::load_all(path = '~/Documents/GitHub/fldgen')

knitr::opts_chunk$set(echo = TRUE)
```


# Setup {.tabset}

## Select model + helpful strings and other files
```{r names}
esmname <- 'GFDL-CM4_ssp245'

outfileyrs <- 2000:2100

# CMIP6 offers gr (regridded) and gn (native) resolutions; cannot get gn to open 
# in some cases, there are gr1, gr2, etc. Unclear the details/differences. 
# probably regridded to different scale. Fine as long as consistent within
# a given variable (tos vs tas)
gridchoice <- 'gr' 


coralsites <- as_tibble(read.csv('../data/cell_site_lat_longs.csv'))

hawaii <- paste0(rep('hi', 15), 1:15)
florida <- paste0(rep('fl', 9), 1:9)
pr <- paste0(rep('pr_usvi', 8), 1:8)
```


# Load the pattern
```{r pattern load}

# the pattern is only for the ocean grid cells.
# To make the .OUT files for combo, the full 
# grid is necessary. 
# It's also convenient  to have the grid 
# lon and lat values over the ocean and their
# corresponding colnumber in the `tos` data 
# that has no NAs.
# So these are included in the rds file for 
# each pattern. 

x <- readRDS(paste0('patterns/', esmname, '_annualtos.rds'))

annual_pattern <- x$annual_pattern
lonlat <- x$lonlat
fulllon <- x$fulllon
fulllat <- x$fulllat

rm(x)
```

# Load and process hector data as needed
Beacuse the ESM `tas` data used to calculate `tgav` and form the pattern is in terms of actual degrees, not anomaly from pre-industrial, Hector temperatures MUST be converted to actual temperature via adding 288.15.


```{r hectorload}

# hector rcp4.5 tgav trajectory
as_tibble(read.csv('hectordata/hector_temp_rcp45.csv', stringsAsFactors = F)) %>%
  # convert from delta tgav to tgav and then to celsius %>%
  mutate(temp = temp + 288.15 - 273.15) ->
  hectortgav


# Hector rcp26 CO2 trajectory
as_tibble(read.csv('hectordata/sample_outputstream_rcp26.csv', stringsAsFactors = F, comment.char = '#')) %>%
     filter(variable == 'Ca',
            spinup == 0, 
            year >= 2000,
            year <= 2100)  %>% select(year, value) %>%
     rename(YEAR = year, CO2 = value) %>%
  write.table(., 'hectordata/CO2conc_hector26.txt', sep = '\t',
              quote = FALSE, row.names=FALSE)


# hector rcp2.6 tgav trajectory
as_tibble(read.csv('hectordata/sample_outputstream_rcp26.csv', stringsAsFactors = F, comment.char = '#')) %>%
  filter(variable == 'Tgav',
         spinup == 0, 
         year >= 2000,
         year <= 2100) %>%
  rename(temp = value) %>%
  # convert from delta tgav to tgav and then to celsius %>%
  mutate(temp = temp + 288.15 - 273.15) ->
  hectortgav26

write.csv(hectortgav26, 'hectordata/hector_temp_rcp26.csv', row.names = F)
```

# Construct the out files

## Hector rcp4.5 with the pattern


```{r}

# Get the annual TOS in each grid cell as predicted 
# by the annual pattern for the Hector tgav
hector_annual_tos <- pscl_apply(annual_pattern, as.vector(hectortgav$temp))


# helper function to pull off, shape, and save  a year's data
make_single_year_output <- function(year){
  
  ind <- which(hectortgav$year == year)
  
  tibble(tos = hector_annual_tos[ind,]) %>%
  mutate(colnumber = as.numeric(row.names(.))) %>%
  left_join(lonlat, by = 'colnumber') %>%
  select(-colnumber) %>% 
  mutate(tos = round(tos, digits = 2)) %>%
  complete(gridlat = fulllat, gridlon = fulllon) %>%
  replace_na(list(tos = -999)) %>%
  spread(gridlon, tos) %>%
  arrange(-gridlat) %>%
  # numeric column names so matlab understands
  rename( `1` = gridlat) %>% 
  mutate(`2` = `1`) -> 
  # rename( `degN/degE` = gridlat) %>%
  # mutate(`: degN` = `degN/degE`) -> 
  hector_annual_tos_year
  
  hector_annual_tos_year <- rbind(hector_annual_tos_year, names(hector_annual_tos_year))
 
  # csv with numeric column names and no header helps matlab
  write.table(hector_annual_tos_year,
            paste0('pscl_out_files/esmgrid/', year, '_', esmname, '.csv'),
            sep = ",", quote = FALSE, row.names=FALSE)

  return(hector_annual_tos_year)

}


# make the outfile for every year
invisible(sapply(outfileyrs, make_single_year_output))
```


## Hector rcp2.6 with the pattern
```{r}



# by the annual pattern for the Hector tgav
hector_annual_tos26 <- pscl_apply(annual_pattern, as.vector(hectortgav26$temp))


make_single_year_output26 <- function(year){
  
ind <- which(hectortgav26$year == year)

tibble(tos = hector_annual_tos26[ind,]) %>%
  mutate(colnumber = as.numeric(row.names(.))) %>%
  left_join(lonlat, by = 'colnumber') %>%
  select(-colnumber) %>% 
  mutate(tos = round(tos, digits = 2)) %>%
  complete(gridlat = fulllat, gridlon = fulllon) %>%
  replace_na(list(tos = -999)) %>%
  spread(gridlon, tos) %>%
  arrange(-gridlat) %>%
  # numeric column names so matlab understands
  rename( `1` = gridlat) %>% 
  mutate(`2` = `1`) -> 
  hector_annual_tos26_year

hector_annual_tos26_year <- rbind(hector_annual_tos26_year, names(hector_annual_tos26_year))

# csv with numeric column names and no header helps matlab
write.table(hector_annual_tos26_year,
            paste0('pscl_out_files/esmgrid/', year,'_', esmname, '_hector26.csv'),
            sep = ",", quote = FALSE, row.names=FALSE)

return(hector_annual_tos26_year)
}


# make the outfile for every year
invisible(sapply(outfileyrs, make_single_year_output26))
```

